---
title: "W12-L2-Lecture-Examples"
author: "Dr. Coleman"
output: html_document
---

## Example: Linear Model

```{r message=FALSE}

library(tidyverse)
library(broom)

fit <- lm(Ozone ~ Wind + Temp + Solar.R,
          data = airquality)

tidy(fit)
```

Check relationships

```{r}

airquality |> 
  ggplot(aes(x = Wind, y = Ozone)) +
  geom_point()
```

Natural spline

```{r}

library(splines)
fit <- lm(Ozone ~ ns(Wind, 2) + Temp + Solar.R,
          data = airquality)

tidy(fit)
```

Data preprocessing

```{r}

aq2 <- airquality |> 
    mutate(Temp = scale(Temp),
           Wind = scale(Wind),
           Solar.R = scale(Solar.R))

fit <- lm(Ozone ~ ns(Wind, 2) + Temp + Solar.R, 
          data = aq2)

tidy(fit)
```

## Tidymodels: Linear Regression

```{r message=FALSE}

library(tidymodels)
rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R) 

model <- linear_reg() |> 
    set_engine("lm") |> 
    set_mode("regression")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

model_fit <- fit(wf, data = airquality)

model_fit

tidy(model_fit)
```

Natural spline

```{r}

rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R) |> 
    step_ns(Wind, deg_free = 2)

model <- linear_reg() |> 
    set_engine("lm") |> 
    set_mode("regression")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

model_fit <- fit(wf, data = airquality)

model_fit

tidy(model_fit)
```

Normalize

```{r}

rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R) |> 
    step_normalize(Wind, Temp, Solar.R) |> 
    step_ns(Wind, deg_free = 2)

model <- linear_reg() |> 
    set_engine("lm") |> 
    set_mode("regression")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

model_fit <- fit(wf, data = airquality)

model_fit

tidy(model_fit)
```

## Check model fit

```{r}

rec |> 
    prep(airquality)

dat_model <- rec |> 
    prep(airquality) |> 
    bake(new_data = NULL)

dat_model

airquality |> 
  tibble() 

model_fit |> 
    extract_fit_parsnip() |> 
    augment(new_data = dat_model)

model_fit |> 
    extract_fit_parsnip() |> 
    augment(new_data = dat_model) |> 
    ggplot(aes(.pred, Ozone)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1)
```

## Tidymodels: Logistic Regression

```{r}

rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R) |> 
    step_naomit(Ozone) |> 
    step_cut(Ozone, breaks = 70) |> 
    step_normalize(Wind, Temp, Solar.R) |> 
    step_ns(Wind, deg_free = 2)

model <- logistic_reg() |> 
    set_engine("glm") |> 
    set_mode("classification")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

model_fit <- fit(wf, data = airquality)

tidy(model_fit)
```

Checking model fit

```{r}

dat_model <- rec |> 
    prep(airquality) |> 
    bake(new_data = NULL)

model_fit |> 
    extract_fit_parsnip() |> 
    augment(new_data = dat_model)

model_fit |> 
    extract_fit_parsnip() |> 
    augment(new_data = dat_model) |> 
    ggplot(aes(`.pred_(70,168]`, Ozone)) + 
    geom_point(alpha = 1/2) +
    labs(x = "Predicted Probability Ozone > 70 ppb",
         y = "Ozone Classification")
```

## Estimating Model Performance on Training Data

### Linear Regression

```{r}

rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R)

model <- linear_reg() |> 
    set_engine("lm") |> 
    set_mode("regression")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

folds <- vfold_cv(airquality, v = 5)

folds

model_fit <- fit_resamples(wf, resamples = folds)

model_fit

model_fit |> 
    collect_metrics()
```

### Logistic Regression

```{r}

rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R) |> 
    step_naomit(Ozone) |> 
    step_cut(Ozone, breaks = 70)

model <- logistic_reg() |> 
    set_engine("glm") |> 
    set_mode("classification")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

folds <- vfold_cv(airquality, v = 5)

folds

model_fit <- fit_resamples(wf, resamples = folds)

model_fit |> 
    collect_metrics()
```

Fit on complete dataset

```{r}

model_fit <- fit(wf, data = airquality)

dat_model <- rec |> 
    prep(airquality) |> 
    bake(new_data = NULL)

a <- model_fit |> 
    extract_fit_parsnip() |> 
    augment(new_data = dat_model)

table(predicted = a$.pred_class, 
      truth = a$Ozone)
```

## Changing models

Original approach

```{r}

rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R) |> 
    step_naomit(everything()) |> 
    step_cut(Ozone, breaks = 70)

model <- logistic_reg() |> 
    set_engine("glm") |> 
    set_mode("classification")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

folds <- vfold_cv(airquality, v = 5)
folds

model_fit <- fit_resamples(wf, resamples = folds)

model_fit |> 
    collect_metrics()
```

Decision tree approach

```{r}

model <- decision_tree() |> 
    set_engine("rpart") |> 
    set_mode("classification")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

folds <- vfold_cv(airquality, v = 5)
folds

model_fit <- fit_resamples(wf, resamples = folds)

model_fit |> 
    collect_metrics()
```

## Tuning parameters

```{r}

rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R) |> 
    step_ns(Wind, deg_free = 2)

model <- linear_reg() |> 
    set_engine("lm") |> 
    set_mode("regression")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

model_fit <- fit(wf, data = airquality)

tidy(model_fit)
```

Adding in cross-validation

```{r}

rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R) |> 
    step_ns(Wind, deg_free = tune("df"))

model <- linear_reg() |> 
    set_engine("lm") |> 
    set_mode("regression")

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

folds <- vfold_cv(airquality, v = 5)

res <- tune_grid(wf, resamples = folds, 
                 grid = tibble(df = c(2, 4, 6, 8)))

res
```

Collect metrics

```{r}

res |> 
  collect_metrics()
```

Find lowest rmse

```{r}

res |> 
    collect_metrics() |> 
    filter(.metric == "rmse") |> 
    arrange(mean)
```

Find highest R squared

```{r}

res |> 
    collect_metrics() |> 
    filter(.metric == "rsq") |> 
    arrange(desc(mean))

```

## Multiple Tuning Parameters

```{r}

rec <- airquality |> 
    recipe(Ozone ~ Wind + Temp + Solar.R) |> 
    step_normalize(Wind, Temp, Solar.R) |> 
    step_ns(Wind, deg_free = tune("df_wind")) |> 
    step_ns(Temp, deg_free = tune("df_temp"))

wf <- workflow() |> 
    add_recipe(rec) |> 
    add_model(model)

res <- tune_grid(wf, resamples = folds, 
                 grid = expand.grid(df_wind = c(2, 4, 6, 8),
                                    df_temp = c(2, 4, 6, 8)),
                 control = control_grid(save_pred = TRUE))

expand.grid(df_wind = c(2, 4, 6, 8),
            df_temp = c(2, 4, 6, 8))

res |> 
    collect_metrics() |> 
    filter(.metric == "rmse") |> 
    arrange(mean)
```

Show best values

```{r}

res |> 
    show_best(metric = "rmse")
```

Select absolute best

```{r}

res |> 
    select_best(metric = "rmse")
```
